rbeast2: BEAUti 2, BEAST2 and Tracer for R

### 2018-02-16 article.tex
Feedback @Giappo all in

  \textbf{1. }
    In the field of phylogenetics, 
     BEAST2 is one of the most widely used software tools. 
     It comes with the graphical programs BEAUti 2 to
     facilitate the creation of BEAST2 configuration files, and
     with DensiTree and Tracer, to interpret BEAST2's output files. 
     However, when many different alignments or model 
     setups are required, a workflow of GUI programs is cumbersome. \\
  \textbf{2. }
    Here, we present a free, libre and open-source package, \verb;rbeast2;, 
    'BEAUti 2, BEAST2 and Tracer for R', for the R programming language. 
    \verb;rbeast2; creates BEAST2 input files, runs BEAST2 and parses its results, 
    all from an R function call. \\
  \textbf{3. }
    We describe \verb;rbeast2;'s usage, the novel functionality it provides
    compared to the tools it is alternative of, and give some examples. \\
  \textbf{4. }
    As \verb;rbeast2; is designed to be of high quality and extendable, 
    we conclude by describing the further development of the package \\
\end{abstract}

@@

There are many phylogenetics tools available to obtain an estimate 
of the phylogenetic tree of a given set of species. 
BEAST2 \cite{bouckaert2014beast} is one of the most widely used ones. 
It creates a posterior of jointly-estimated phylogenies and model parameters, 
from a DNA, RNA or amino acid alignment (see figure \ref{fig:workflow} 
for an overview of the workflow). 
It is a console application, that needs a configuration file containing 
alignments and model parameters.

@@

BEAUti cannot be called from a command-line script.
This implies that when the user 
wants to explore the consequences of various settings, this must be done manually.
This is the common workflow when using a few alignments and doing a superficial 
analysis of sensitivity of the reconstructed tree to model settings.

@@

However, for exploring the output of many BEAST2 runs, 
one would like a script to collect all parameter ESSes,
parameter traces and posterior phylogenies.

Here, to provide such functionality we present \verb;rbeast2;, 
’BEAUti 2, BEAST2 and Tracer for R’, 
which creates BEAST2 configuration files,
runs BEAST2, and analyzes its results,
all from an R function call. This
will save time, tedious mouse clicking and 
reduces the chances of errors in such repetitive actions.
The interface of \verb;rbeast2; mimics the tools it
is inspired on. This
familiarity helps both beginner and experienced BEAST2 users 
to make the step from those tools to \verb;rbeast2;.
\verb;rbeast2; enables the creation of a single-script 
pipeline from sequence alignments to posterior analysis in R.

\verb;rbeast2; is the first R package that unifies the
full workflow of working with BEAST2. An
example to create a BEAST2 input file is 
\verb;BEASTmasteR; \cite{beastmaster}.
Also \verb;BEASTmasteR; allows 
to create BEAST2 configuration files from R. 
The difference is that \verb;rbeast2; 
has its focus on DNA alignments and ultrametric trees, 
where \verb;BEASTmasteR; is used for 
morphological traits and tip-dating.
Examples of R packages to parse the BEAST2 output files
are \verb;rBEAST; \cite{rBEAST} and 
\verb;RBEAST; \cite{RBEAST}.

>>>>

  \textbf{1. }
    In the field of phylogenetics, 
    BEAST2 is one of the most widely used software tools. 
    It comes with the graphical programs BEAUti 2, DensiTree and Tracer,
    to create BEAST2 configuration files and to interpret BEAST2's output files. 
    However, when many different alignments or model 
    setups are required, a workflow of GUI programs is cumbersome. \\
  \textbf{2. }
    Here, we present a free, libre and open-source package, \verb;rbeast2;: 
    'BEAUti 2, BEAST2 and Tracer for R', for the R programming language. 
    \verb;rbeast2; creates BEAST2 input files, runs BEAST2 and parses its results, 
    all from an R function call. \\
  \textbf{3. }
    We describe \verb;rbeast2;'s usage, the novel functionality it provides
    compared to the tools it is alternative of, and give some examples. \\
  \textbf{4. }
    As \verb;rbeast2; is designed to be of high quality and extendable, 
    we conclude by describing the further development of the package. \\
\end{abstract}

@@

There are many phylogenetics tools available to obtain an estimate 
of the phylogenetic tree of a given set of species. 
BEAST2 \cite{bouckaert2014beast} is one of the most widely used ones. 
It creates a posterior of jointly-estimated phylogenies and model parameters, 
from one or more DNA, RNA or amino acid alignments (see figure \ref{fig:workflow} 
for an overview of the workflow). 
It is a console application, that needs a configuration file containing 
alignments and model parameters.

@@

However, BEAUti cannot be called from a command-line script.
This implies that when the user 
wants to explore the consequences of various settings, this must be done manually.
This is the common workflow when using a few alignments and doing a superficial 
analysis of sensitivity of the reconstructed tree to model settings.

@@

However, for exploring the output of many BEAST2 runs, 
one would like a script to collect all parameters' ESSes,
parameter traces and posterior phylogenies.

Here, to provide such functionality we present \verb;rbeast2;:
’BEAUti 2, BEAST2 and Tracer for R’, 
which creates BEAST2 configuration files,
runs BEAST2, and analyzes its results,
all from an R function call. This
will save time, tedious mouse clicking and 
reduces the chances of errors in such repetitive actions.
The interface of \verb;rbeast2; mimics the tools it
is inspired on. This
familiarity helps both beginner and experienced BEAST2 users 
to make the step from those tools to \verb;rbeast2;.
\verb;rbeast2; enables the creation of a single-script 
pipeline from sequence alignments to posterior analysis in R.

\verb;rbeast2; is the first R package that unifies the
full workflow of working with BEAST2. An
example to create a BEAST2 input file is 
\verb;BEASTmasteR; \cite{beastmaster}.
Also \verb;BEASTmasteR; allows 
to create BEAST2 configuration files from R. 
The difference is that \verb;rbeast2; 
has its focus on DNA alignments and ultrametric trees (that
is, trees that have their tips measured in the present), 
where \verb;BEASTmasteR; is used for 
morphological traits and tip-dating.
Examples of R packages to parse the BEAST2 output files
are \verb;rBEAST; \cite{rBEAST} and 
\verb;RBEAST; \cite{RBEAST}.
